---
title: "Assignment 3 Task 4 - Mari Herbst de Cortina"
author: "Mari Herbst"
date: "10/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(lubridate)
```

# Naples Reef Urchins

data citation: Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

This data measures red and purple urchin populations and diameter (to nearest 0.5 cm) at 5 sites between 2008 and 2020. The data collection related to urchin populations where giant kelp is left intact vs where it is removed once annually.

### Read in the data using here() to create a system agnostic file path

```{r}
urchins <- read_csv(here("data", "urchins.csv"))
```

- First, I want to make the data more coder friendly.

```{r}
clean_urchins <- clean_names(urchins)
```

## Create a subset of clean_urchins containing:

- Only for collections from the year 2010 (first, change class of date column to "date" using lubridate::mdy)
- Only purple urchins
- Only urchin data collected at Naples Reef

```{r}
napl_urchins <- clean_urchins %>%
  mutate(date = mdy(date)) %>% # change class of column
  mutate(year = year(date)) %>% # create column for year only
  filter(year == 2010) %>% # filter to only 2010
  filter(common_name == "Purple Urchin") %>% # only purple urchins
  filter(site == "NAPL") # only at Naples Reef
```

## Create a faceted histogram and q-q plot for the size distribution of urchins depending on treatment

```{r}
ggplot(data = napl_urchins, aes(x = size)) + # select data and axes
  geom_histogram(bin = 10, fill = "darkolivegreen") + # select graph type and number of bins, color bins
  facet_wrap(~ treatment) + # break graph into one for each treatment
  labs(x = "diameter (cm)") # name x axis

ggplot(data = napl_urchins, aes(sample = size)) + # select data. q-q requires sample = variable
  geom_qq() + # select graph type 
  facet_wrap(~ treatment) + # break plot into one for each treatment
  theme_dark() # stylistic choice
```

Main takeaways: The samples from populations undergoing both treatments have relatively normally distributed urchin diameter/size. Smaller urchins seem to have more success in areas where kelp was cleared annually, whereas areas where kelp was not removed had a more even distribution of sizes.
