---
title: "Assignment 3 Task 3 - Mari Herbst de Cortina"
author: "Mari Herbst"
date: "10/24/2021"
output: html_document
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(here)
library(tidyverse)
library(dplyr)
library(janitor)
```

### Read in the data using here() to create a system agnostic file path
```{r}
urchins <- read_csv(here("data", "urchins.csv"))
```

### Section 1: size comparisons for purple and red urchins

- First, I want to make the data more coder friendly

```{r}
clean_urchins <- clean_names(urchins)
```


#### Data visualization - histogram

- Line 1: Select dataframe and the variable to be counted
- Line 2: Choose the type of graph, in this case, we are making a histogram. Also, select the variable by which to color code 
- Line 3-7: Use labs() to name axes, title, and legend title
- Line 8: Use facet_wrap() to divide the graph into 2 histograms, one for each common name of urchins

```{r}
ggplot(data = clean_urchins, aes(x = size)) +
  geom_histogram(bins = 10, aes(fill = common_name)) +
  labs(x = "urchin size (cm)",
       title = "Distribution of Purple vs. Red Urchin Size",
       legend = "Urchin species",
       fill = "Urchin type") +
  facet_wrap(vars(common_name))
```



##### Figure 1: Distribution of Purple vs. Red Urchin Size: 50 red and 50 purple urchins were measured for diameter (shell without spines) to the closest half centimeter across 5 sites on California's coast between 2008 and 2011


