---
title: "Assignment 3 Task 3 - Mari Herbst de Cortina"
author: "Mari Herbst"
date: "10/24/2021"
output: html_document
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(here)
library(tidyverse)
library(dplyr)
library(janitor)
library(kableExtra)
library(lubridate)
```

# Exploring Urchins

#### data citation: Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

#### This data measures red and purple urchin populations and diameter (to nearest 0.5 cm) at 5 sites between 2008 and 2020. The data collection related to urchin populations where giant kelp is left intact vs where it is removed once annually.

### Read in the data using here() to create a system agnostic file path

```{r}
urchins <- read_csv(here("data", "urchins.csv"))
```

## Section 1: size comparisons for purple and red urchins

- First, I want to make the data more coder friendly.

```{r}
clean_urchins <- clean_names(urchins)
```


### Data visualization - histogram

- Select dataframe and the variable to be counted.
- Choose the type of graph, in this case, we are making a histogram. Also, select the variable by which to color code and select the number of bins for the histogram to organize data into.
- Use labs() to name axes, title, and legend title.
- Use facet_wrap() to divide the graph into 2 histograms, one for each common name of urchins.

```{r}
ggplot(data = clean_urchins, aes(x = size)) +
  geom_histogram(bins = 10, aes(fill = common_name)) +
  labs(x = "urchin diameter (cm)",
       title = "Distribution of Purple vs. Red Urchin Size",
       legend = "Urchin species",
       fill = "Urchin type") +
  facet_wrap(vars(common_name))
```



##### Figure 1. Distribution of Purple vs. Red Urchin Size: Red and purple urchins were measured for diameter (shell without spines) to the closest half centimeter across 5 sites on California's coast between 2008 and 2011.

### Descriptive statistics table - summary table

- Create a summary table grouped by urchin type that includes the mean and median size, the standard deviation in size, and the sample size for each urchin type.
- Be sure to use kable() to name columns appropriately and with units!

```{r}
urchins_summary <- clean_urchins %>%
  group_by(common_name) %>%
  summarize(mean_size_cm = round(mean(size, na.rm = TRUE), 2),
            median_size_cm = round(median(size), 2),
            sd_size = round(sd(size), 2),
            sample_size = n())

urchins_summary %>%
  kable(col.names = c("Species",
                      "Mean diameter (cm)",
                      "Median diameter (cm)",
                      "Standard deviation (cm)",
                      "Sample Size"),
        caption = "Table 1: A summary of urchin sizes showing the mean and median size, standard deviation in size, and sample size of each type of urchin that was studied.")
```

##### In general, red urchins are larger than purple urchins in this study area; the average purple urchin is 2.31cm smaller than the average red one. Red urchins are, on average, 1.52 times as large as purple urchins. There is a greater degree of spread amongst red urchins, indicated by the higher standard deviation figure.


## Section 2: urchin counts by site and year

### Create a subset of data

- only include purple urchin observations using filter()
- convert date class to "Date" using myd()
- create a column just for year using separate()
- create a column called "site_full" containing the written out name of each site using case_when()

```{r}
purple_urchins <- clean_urchins %>%
  filter(common_name == "Purple Urchin") %>%
  mutate(date = mdy(date)) %>%
  separate(col = date,
           into = c("year", "month_day"),
           sep = "-",
           extra = "merge",
           remove = TRUE) %>%
  mutate(site_full = case_when(
    site %in% c("NAPL") ~ "Naples Reef",
    site %in% c("AQUE") ~ "Arroyo Quemado Reef",
    site %in% c("CARP") ~ "Carpinteria Reef",
    site %in% c("MOHK") ~ "Mohawk Reef",
    site %in% c("IVEE") ~ "Isla Vista Reef"))
```

### Create a table containing the total number of purple urchins observed by site and year

- Use group_by() to select site_full and year
- Summarize by sample size using n()

```{r}
purple_urchin_count <- purple_urchins %>%
  group_by(site_full, year) %>%
  summarize(total_urchins = n())
```

### Create a graph of purple_urchin_count site

- Use ggplot2::geom_histogram to make a histogram of urchin counts
- Use labs() to label axes and title
- Use facet_wrap() to break graphs up by site

```{r}
ggplot(data = purple_urchin_count, aes(x = year, y = total_urchins)) +
  geom_col() +
  labs(x = "year",
       y = "purple urchin count",
       title = "Purple Urchin Counts Per Year Per Site") +
  theme(axis.text.x = element_text(angle = 90)) + # change angle of x axis values to make easier to read
  facet_wrap(~ site_full)
```

##### Figure 2. Purple Urchin Counts Per Year Per Site: At all sites aside from Isla Vista Reef, purple urchin counts decreased between 2009 and 2012 and then have stayed relatively consistent through 2020. At Isla Vista Reef, no purple urchins were observed until 2012 and have had a relatively consistent population since then. In order to understand whether these changes are representative of the true urchin populations in these locations, we would need to examine some summary statistics for these samples. The treatments from the study could be affecting urchin populations, as could predation, competition, or weather events

## END TASK 3
